Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC

### 2016-03-28 tese2/14_implementation_signaling.tex
[ADD] performance tests, updated playFW, websocket simplification

Our signaling protocol consists in send and receiving JSON formated messages over WebSockets by both the server and the client.

\begin{lstlisting}[caption={General structure of our WebSocket messages},language=json]
{
	"cmd":<cmd>,
	"data":<data>
}
\end{lstlisting}

@@

If the user decides to share either from camera of screen, \emph{getUserMedia} is called with the correspondent contraints in order to obtain a local stream.

\begin{lstlisting}[caption={Media constraints},language=JavaScript]
var screenShareConstraints = {	
	"video": {
		"mediaSource": "window" || "screen"
	}, 
	"audio": false
};
var cameraMicrophoneConstraints = {
	"audio":true, 
	"video":true 
};
var receiveOnlyConstraints = {
	"offerToReceiveAudio":true,
	"offerToReceiveVideo":true
};
\end{lstlisting}

A popup is raised in order to ask the user to give permission to share those resources. If the resources are shared successfully the user creates an offer, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through WebSockets. If the resources are not shared or the user specified to receive stream only, an offer is created specifying the contraints for receiving only video and audio.

@@

The server receive and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint. Then a \emph{local session description} is created on the server and sent back to the client, after that the server trys to gather \ac{ICE} candidates.

@@

After a while both the server and client receive the \ac{ICE} candidates that allow the client to connect directly to \ac{KMS} and vice-versa. The candidates are received at the client which sets them to its \emph{PeerConnection}. The same is done on the server which receives the \ac{ICE} candidates from the client and propagates them to \ac{KMS}.

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}{}
    	\includegraphics[width=1.0\textwidth]{figures/signaling}
    \end{subfigure}
    \caption{Signaling sequence diagram}
\end{figure}

>>>>

Our signaling protocol consists in send and receiving \ac{JSON} formated messages like listing \ref{lst:msgtype} over WebSockets by both the server and the client.

\begin{minipage}{\linewidth}
\begin{lstlisting}[caption={General structure of our WebSocket messages},label={lst:msgtype},language=json]
{
	"cmd":<cmd>,
	"data":<data>
}
\end{lstlisting}
\end{minipage}

@@

If the user decides to share either from camera of screen, \emph{getUserMedia} is called with the correspondent contraints in order to obtain a local stream. Use the constraints presented on listing \ref{lst:constraints}.

\begin{minipage}{\linewidth}
\begin{lstlisting}[caption={Media constraints},label={lst:constraints},language=JavaScript]
var screenShareConstraints = {	
	"video": {
		"mediaSource": "window" || "screen"
	}, 
	"audio": false
};
var cameraMicrophoneConstraints = {
	"audio":true, 
	"video":true 
};
var receiveOnlyConstraints = {
	"offerToReceiveAudio":true,
	"offerToReceiveVideo":true
};
\end{lstlisting}
\end{minipage}

A popup is raised in order to ask the user to give permission to share those resources. If the resources are shared successfully the user creates an offer like listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through WebSockets. If the resources are not shared or the user specified to receive stream only, an offer is created specifying the contraints for receiving only video and audio.

\begin{minipage}{\linewidth}
\begin{lstlisting}[caption={Offer created by client},label={lst:sig01},language=json]
{
	"cmd":"offer",
	"data":{
		"type":"offer",
		"sdp":<sdp>	// this is huge
	}
}
\end{lstlisting}
\end{minipage}

@@

The server receive and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint. Then a \emph{local session description} like listing \ref{lst:sig02} is created on the server and sent back to the client, after that the server trys to gather \ac{ICE} candidates.

\begin{minipage}{\linewidth}
\begin{lstlisting}[caption={Answer created by KMS},label={lst:sig02},language=json]
{
	"cmd":"answer",
	"data":{
		"type":"answer",
		"sdp":<sdp>	// this is huge
	}
}
\end{lstlisting}
\end{minipage}

@@

After a while both the server and client receive the \ac{ICE} candidates that allow the client to connect directly to \ac{KMS} and vice-versa. The candidates are received at the client which sets them to its \emph{PeerConnection}. The same is done on the server which receives the \ac{ICE} candidates like on listing \ref{lst:sig03} from the client and propagates them to \ac{KMS}.

\begin{minipage}{\linewidth}
\begin{lstlisting}[caption={ICE candidates sent by KMS and client},label={lst:sig03},language=json]
{
	"cmd":"iceCandidate",
	"data":{
		"sdpMLineIndex":0,
		"candidate":"candidate:15 1 TCP 843056127 146.193.224.20 48828 typ srflx raddr 192.168.1.105 rport 48828 tcptype passive",
		"sdpMid":"audio"
	},
}
\end{lstlisting}
\end{minipage}

@@

\begin{figure}[!htb]
    \centering
    \begin{subfigure}{}
    	\includegraphics[width=0.8\textwidth]{figures/signaling}
    \end{subfigure}
    \caption{Signaling sequence diagram}
\end{figure}

### 2016-03-28 tese2/16_implementation_content.tex
[ADD] performance tests, updated playFW, websocket simplification

>>>>

	Listing \ref{lst:pseudo_render} presents the pseudo code for our content scheduler.

\begin{minipage}{\linewidth}
\begin{lstlisting}[caption={Pseudo code for hyper content scheduler},label={lst:pseudo_render},language=JavaScript]
function scheduleContent() {
    var navigatedTime = getCurrentTime() - getUserOffset();
    while ( hasLocalEvents() ) {
        if ( firstEventIsOlderThanNavigatedTime() ) {
            if ( firstEventIsStart() ) {
            	addHtmlLayer(event);
            } else {
            	removeHtmlLayer(event);
            }
            removeFirstEvent();
        } else {
            break;
        }
    }
    if ( hasNoLocalEvents() ) {
    	return;	// nothing to do
    } 
    if ( hasNoStartEvents() && serverHasMoreContent() ) { 
       	requestContentFromServer(navigatedTime);
    } else {
        waitUntilNextEvent();
        scheduleContent();
    }
}
\end{lstlisting}
\end{minipage}

### 2016-03-28 tese2/17_implementation_deploy.tex
[ADD] performance tests, updated playFW, websocket simplification

	\begin{table}[H]
\centering
\caption{Hardware specifications}
\label{table:hwspecs}
\begin{tabular}{|l|l|}
\hline
Server & Supermicro SuperServer 6027R-72RF\footnotemark \\ \hline
CPU & 2 x Intel Xeon E5-2640V2, LGA 2011, 2.0GHz, 8C/16T \\ \hline
RAM & 8 x DDR3 REG16G-1600DDR3, 16GB, DDR3-1600, Registered ECC, memory \\ \hline
Network cards & 2 x Intel Corporation I350 Gigabit Network Connection (rev 01)  \\ \hline
Disks & \begin{tabular}[c]{@{}l@{}}2x SAMSUNG SSD 840 PRO 256GB SATA III (Drive 0 - RAID 1 - 237.486 GB)\\
4x WESTERN DIGITAL 3TB SATA III 64MB RED (Drive 1 - RAID 5 - 8.185TB)\end{tabular} \\ \hline
\end{tabular}
\end{table}

@@

	\begin{table}[H]
\centering
\caption{Operating system specifications}
\label{table:osspecs}
\begin{tabular}{|l|l|}
\hline
Operating System & Linux version 3.16.0-4-amd64 \\ \hline
Distribution & Debian GNU/Linux 8 (jessie)\footnotemark \\ \hline
Swap & 16GB (Drive 0)\\ \hline
Root & 40GB (Drive0)\\ \hline
EFI & 200MB (Drive 0)\\ \hline
Bcache & Remaining space (Drive 0 \& Drive 1) \\ \hline
\end{tabular}
\end{table}

@@

	\begin{table}[H]
\centering
\caption{Installed Software}
\label{table:softspecs}
\begin{tabular}{|l|l|}
\hline
\multicolumn{1}{|c|}{\textbf{Name}} & \multicolumn{1}{c|}{\textbf{Version}}         \\ \hline
Ubuntu Server\footnotemark & 14.04 LTS   \\ \hline
Oracle Java & 1.8.0\_77   \\ \hline
MongoDB & 3.0.10   \\ \hline
Kurento Media Server & 6.4.0 \\ \hline
Kurento Repository & 6.3.1 \\ \hline
Python & 2.7.6 \\ \hline
\end{tabular}
\end{table}
 \footnotetext{\url{http://www.ubuntu.com/} (Accessed March 27, 2016)}

>>>>

	\begin{table}[!htb]
\centering
\caption{Hardware specifications}
\label{table:hwspecs}
\begin{tabular}{|l|l|}
\hline
Server & Supermicro SuperServer 6027R-72RF\footnotemark \\ \hline
CPU & 2 x Intel Xeon E5-2640V2, LGA 2011, 2.0GHz, 8C/16T \\ \hline
RAM & 8 x DDR3 REG16G-1600DDR3, 16GB, DDR3-1600, Registered ECC, memory \\ \hline
Network cards & 2 x Intel Corporation I350 Gigabit Network Connection (rev 01)  \\ \hline
Disks & \begin{tabular}[c]{@{}l@{}}2x SAMSUNG SSD 840 PRO 256GB SATA III (Drive 0 - RAID 1 - 237.486 GB)\\
4x WESTERN DIGITAL 3TB SATA III 64MB RED (Drive 1 - RAID 5 - 8.185TB)\end{tabular} \\ \hline
\end{tabular}
\end{table}

@@

	\begin{table}[!htb]
\centering
\caption{Operating system specifications}
\label{table:osspecs}
\begin{tabular}{|l|l|}
\hline
Operating System & Linux version 3.16.0-4-amd64 \\ \hline
Distribution & Debian GNU/Linux 8 (jessie)\footnotemark \\ \hline
Swap & 16GB (Drive 0)\\ \hline
Root & 40GB (Drive0)\\ \hline
EFI & 200MB (Drive 0)\\ \hline
Bcache & Remaining space (Drive 0 \& Drive 1) \\ \hline
\end{tabular}
\end{table}

@@

	\begin{table}[!htb]
\centering
\caption{Installed Software}
\label{table:softspecs}
\begin{tabular}{|l|l|}
\hline
\multicolumn{1}{|c|}{\textbf{Name}} & \multicolumn{1}{c|}{\textbf{Version}}         \\ \hline
Ubuntu Server\footnotemark & 14.04 LTS   \\ \hline
Oracle Java & 1.8.0\_77   \\ \hline
Play Framework & 2.5.0  \\ \hline
MongoDB & 3.0.10   \\ \hline
Kurento Media Server & 6.4.0 \\ \hline
Kurento Repository & 6.3.1 \\ \hline
Python & 2.7.6 \\ \hline
\end{tabular}
\end{table}
 \footnotetext{\url{http://www.ubuntu.com/} (Accessed March 27, 2016)}

### 2016-03-28 tese2/evaluation.tex
[ADD] performance tests, updated playFW, websocket simplification

  \begin{table}[H]
\centering
\caption{Mapping of connections by network interface}
\label{table:interfacemap}
\begin{tabular}{|l|l|}
\hline
\multicolumn{1}{|c|}{\textbf{Interface}} & \multicolumn{1}{c|}{\textbf{Connection}}         \\ \hline
Loopback & Web Server $\leftrightarrow$ MongoDB  \\ \hline
Loopback & Web Server $\leftrightarrow$ Kurento Media Server  \\ \hline
Loopback & Kurento Media Server $\leftrightarrow$ Kurento Repository $\leftrightarrow$ MongoDB  \\ \hline
Ethernet & Web Server $\leftrightarrow$ Client\\ \hline
Ethernet & Kurento Media Server $\leftrightarrow$ Client\\ \hline
\end{tabular}
\end{table}

@@

\begin{figure}[H]
  \begin{center}
    \input{stats/test_full_features_net.tex}
  \end{center}
  \caption{Network usage after implementing all features}
  \label{fig:test_full_features_net}
\end{figure}

@@

\begin{figure}[H]
  \begin{center}
    \input{stats/summary_full_net.tex}
  \end{center}
  \caption{Average network usage per interval of events after implementing all features}
  \label{fig:summary_full_net}
\end{figure}

@@

>>>>

  \begin{table}[!htb]
\centering
\caption{Mapping of connections by network interface}
\label{table:interfacemap}
\begin{tabular}{|l|l|}
\hline
\multicolumn{1}{|c|}{\textbf{Interface}} & \multicolumn{1}{c|}{\textbf{Connection}}         \\ \hline
Loopback & Web Server $\leftrightarrow$ MongoDB  \\ \hline
Loopback & Web Server $\leftrightarrow$ Kurento Media Server  \\ \hline
Loopback & Kurento Media Server $\leftrightarrow$ Kurento Repository $\leftrightarrow$ MongoDB  \\ \hline
Ethernet & Web Server $\leftrightarrow$ Client\\ \hline
Ethernet & Kurento Media Server $\leftrightarrow$ Client\\ \hline
\end{tabular}
\end{table}

@@

\begin{figure}[!htb]
  \begin{center}
    \input{stats/test_full_features_net.tex}
  \end{center}
  \caption{Network usage after implementing all features}
  \label{fig:test_full_features_net}
\end{figure}

@@

\begin{figure}[!htb]
  \begin{center}
    \input{stats/summary_full_net.tex}
  \end{center}
  \caption{Average network usage per interval of events after implementing all features}
  \label{fig:summary_full_net}
\end{figure}

@@

Figure \ref{fig:test_full_features_mem} shows the memory usage during our performance test case. Both \ac{JVM}, \emph{MongoDB} and \ac{KMS} performs their own memory management by holding and recycling objects when needed. The expected and observed behaviour of the memory usage is growth memory usage while the users are entering in the conference room and a memory usage stabilization aftwerwards.

\begin{figure}[!htb]
  \begin{center}
    \input{stats/test_full_features_mem.tex}
  \end{center}
  \caption{Memory usage after implementing all features}
  \label{fig:test_full_features_mem}
\end{figure}

\emph{MongoDB} memory usage keeps increasing because it trys to fit part of the database on ram for fast read access. \emph{MongoDB} checkpoints data to disk every 60 seconds or when journal data exceeds 2GB\footnote{\url{https://docs.mongodb.org/manual/faq/storage/}(Accessed March 28, 2016)}, that explains the small memory usage peaks during our test case and stabilization at the end.

Figure \ref{fig:test_full_features_cpu} shows the percentage of \ac{CPU} usage during our performance test case. As we can see, the amount of \ac{CPU} increases and decreases linearly in function of the amount of conference participants.

\begin{figure}[!htb]
  \begin{center}
    \input{stats/test_full_features_cpu.tex}
  \end{center}
  \caption{Percentage of CPU used during the performance tests}
  \label{fig:test_full_features_cpu}
\end{figure}

Figure \ref{fig:test_full_features_cpu_zoom} is the same as figure \ref{fig:test_full_features_cpu} but zoomed over \emph{MongoDB}, \emph{Java} and \emph{Kurento Repository}. As we can see there are periodic proccessing peaks every ten seconds that are due to block recording.

\begin{figure}[!htb]
  \begin{center}
    \input{stats/test_full_features_cpu_zoom.tex}
  \end{center}
  \caption{Percentage of CPU used during the performance tests}
  \label{fig:test_full_features_cpu_zoom}
\end{figure}

   Figure \ref{fig:summary_full_cpu} shows the average used \ac{CPU} per interval of consecutive events.

\begin{figure}[!htb]
  \begin{center}
    \input{stats/summary_full_cpu.tex}
  \end{center}
  \caption{Average percentage of CPU used per interval of events after implementing all features (zoomed)}
  \label{fig:summary_full_cpu}
\end{figure}

  Just for testing purposes we have done the same performance tests without detecting \emph{QR codes} in order to understand how \ac{CPU} intensive is this task. Figure \ref{fig:test_without_qrcode_cpu} shows the results for the same experience but with \emph{QR code} detection disabled.

  \begin{figure}[!htb]
  \begin{center}
    \input{stats/test_without_qrcode_cpu.tex}
  \end{center}
  \caption{Percentage of CPU used during the performance tests without QR code detection}
  \label{fig:test_without_qrcode_cpu}
\end{figure}

We conclude that \emph{QR code} detection is a very intensive task, approximately doubling the amount of work performed by \ac{CPU}. Without this feature the network and memory usage had insignificate changes compared to \ac{CPU} usage.

### 2016-03-28 tese2/stats/summary_full_net.tex
[ADD] performance tests, updated playFW, websocket simplification

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=8cm,
	    title={Network usage},
	    xlabel={Time [seconds]},
	    ylabel={Transfer rate [kB/s]},
	    xmin=-10, xmax=790,
	    ymin=0, ymax=2000,
	    xtick={0,60,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

>>>>

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Network usage},
	    xlabel={Time [seconds]},
	    ylabel={Transfer rate [kB/s]},
	    xmin=-10, xmax=850,
	    ymin=0, ymax=1400,
	    xtick={0,60,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

### 2016-03-28 tese2/stats/test_full_features_cpu.tex
[ADD] performance tests, updated playFW, websocket simplification

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Cpu Usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=-10, xmax=790,
	    ymin=0, ymax=1200,
	    xtick={0,60,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

@@

	\end{axis}
	\caption{Percentage of CPU used during the performance tests}
	\label{fig:test_full_features_cpu}
\end{tikzpicture}

>>>>

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={CPU usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=-10, xmax=850,
	    ymin=0, ymax=1200,
	    xtick={0,60,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

@@

	\end{axis}

\end{tikzpicture}

### 2016-03-28 tese2/stats/test_full_features_cpu_zoom.tex
[ADD] performance tests, updated playFW, websocket simplification

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={CPU usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=-10, xmax=850,
	    ymin=0, ymax=1200,
	    xtick={0,60,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

>>>>

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={CPU usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=-10, xmax=850,
	    ymin=0, ymax=100,
	    xtick={0,60,...,2000},
	    ytick={0,20,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

### 2016-03-28 tese2/stats/test_full_features_mem.tex
[ADD] performance tests, updated playFW, websocket simplification

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Mem Usage},
	    xlabel={Time [seconds]},
	    ylabel={Memory [MB]},
	    xmin=-10, xmax=790,
	    ymin=0, ymax=2000,
	    xtick={0,60,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

>>>>

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Memory Usage},
	    xlabel={Time [seconds]},
	    ylabel={Memory [MB]},
	    xmin=-10, xmax=850,
	    ymin=0, ymax=1800,
	    xtick={0,60,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

### 2016-03-28 tese2/stats/test_without_qrcode_mem.tex
[ADD] performance tests, updated playFW, websocket simplification

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Mem Usage},
	    xlabel={Time [seconds]},
	    ylabel={Memory [MB]},
	    xmin=-10, xmax=790,
	    ymin=0, ymax=2000,
	    xtick={0,100,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={10,70,...,1000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

>>>>

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Mem Usage},
	    xlabel={Time [seconds]},
	    ylabel={Memory [MB]},
	    xmin=-10, xmax=850,
	    ymin=0, ymax=2000,
	    xtick={0,100,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={10,70,...,1000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]
